---
title: "serotrackr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{serotrackr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r set_options, include = FALSE}
knitr::opts_chunk$set(
  comment = "#>"
)
```


*serotrackr* helps you prepare your raw data for submission to [SeroTracker](https://serotracker.com){target="_blank"}. I will use parts of a sample raw dataset included in *serotrackr* to demonstrate the workflow of the package:

```{r setup, message=FALSE}
library(serotrackr)
```

```{r subset_sample_raw_data}
mydata <- dplyr::select(
  sample_raw_data,
  dataset_id, id, age_group, age, sex, state, city, result, result_cat
)

mydata
```


Here are the steps you should follow:

## 1) Validate your data

The first step is to use `st_validate()`. It checks your individual-level raw data for common issues, prints a detailed explanation of any errors, and if all is good, outputs a cleaned and validated dataframe. Most of the arguments of this function accept both an unquoted column name or a length-one vector, the value of which will be applied to all records in your data after validation.


```{r st_validate_error}
validated_df <- st_validate(
  sample_raw_data,
  dataset_id = dataset_id,
  id = id,
  age_group = age_group,
  age = age,
  sex = sex,
  adm0 = regions$adm0$Canada,
  adm1 = state,
  adm2 = city,
  collection_start_date = "2020-Mar-01",
  collection_end_date = "15/8/2023",
  test_id = assays$`SARS-CoV-2`$`ID.Vet - IgG - ID Screen`,
  result = result,
  result_cat = result_cat,
  include_others = TRUE,
  rmd_safe = TRUE
)
```

  
<br>
As you can see above, `st_validate()` found 4 issues which must be addressed before proceeding.

## 2) Address issues

### 2.1) General errors

I start with correcting the `age` issues, both of which seem to result from indicating missing values with either `-999` or `999`. In general, any missingness must be represented by `NA` values before running `st_validate()`. Here is a quick fix:

```{r address_age}
mydata <- dplyr::mutate(
  mydata,
  age = ifelse(age %in% c(-999, 999), NA, age)
)
```

### 2.2) Errors in region data

The `adm1` (state/province) and `adm2` (district/municipality) errors happened because `st_validate()` expects standardized region codes, not region names, for these two arguments. These region codes are stored in the `serotrackr::regions` object, which is a named list.

- If you conducted your study in only one region, simply use `regions` to get its region code.
- If you conducted your study in multiple regions and have a column in your data for it, similar to our case here, then use `st_locate()`. It will automatically convert as much of your region names as possible to region codes. The rest must be defined by you, as below.

First, run `st_locate()` with its first four arguments to see what is wrong with your region data:

```{r st_locate_error}
st_locate(data = mydata,
          adm0 = regions$adm0$Canada,
          adm1 = state,
          adm2 = city)
```

Ah, typos! Next, use the ... argument to define the unmatched ones:

```{r st_locate_success}
mydata <- st_locate(
  data = mydata,
  adm0 = regions$adm0$Canada,
  adm1 = state,
  adm2 = city,
  "Toronoto" = regions$adm2$Canada$Ontario$Toronto,
  "Calagry"  = regions$adm2$Canada$Alberta$Calgary
)
```
`st_locate()` added two new columns to the data, named `adm1` and `adm2`, containing standardized region codes. Now that all issues are resolved, let's rerun `st_validate()`:

```{r st_validate_success}
validated_df <- st_validate(
  mydata,
  dataset_id = dataset_id,
  id = id,
  age_group = age_group,
  age = age,
  sex = sex,
  adm0 = regions$adm0$Canada,
  adm1 = adm1,
  adm2 = adm2,
  collection_start_date = "2020-Mar-01",
  collection_end_date = "15/8/2023",
  test_id = assays$`SARS-CoV-2`$`ID.Vet - IgG - ID Screen`,
  result = result,
  result_cat = result_cat,
  include_others = TRUE,
  rmd_safe = TRUE
)
```

  
<br>
Perfect! Here is a glimpse at the validated data:

```{r print_validated_df}
validated_df
```


## 3) Generate aggregate estimates

Now, use `st_aggregate()` to generate aggregated estimates from the validated data:

```{r st_aggregate_success}
estimates <- st_aggregate(validated_df)
```

And here are the generated estimates:

```{r print_estimates}
estimates
```


## 4) Export to Excel for submission

Finally, use `st_save()` to export both validated data and generated estimates to an Excel document:

```{r st_save, eval=FALSE}
st_save(validated_df, estimates, path = "submission.xlsx")
```

The saved Excel document has six sheets, including `Instructions`, `Study Metadata`, `Age Groups`, `Immunoassays`, `Estimates`, and `Data`.

- The output of `st_validate()` fills in the *Age Groups* and *Data* sheets completely. It also fills in parts of the *Study Metadata* and *Immunoassays* sheets, as much as possible.
- The output of `st_aggregate()` fills in the *Estimates* sheet.

Open the Excel document and specifically look for empty mandatory columns, indicated with colour, in *Study Metadata* and *Immunoassays* sheets, and fill them in. When you are done, the document is ready to be submitted on [SeroTracker.com](https://serotracker.com){target="_blank"}.
